---
title: "EcoPractical"
author: "Team 8"
date: "13/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prophet Model 
    • It is a nonlinear regression model 
    • Prophet is a procedure for forecasting time series data based on an additive model where non-linear trends are fit with yearly, weekly, and daily seasonality, plus holiday effects
    • It works best with time series that have strong seasonal effects and several seasons of historical data.
    • It is robust to missing values and shift in trend, handles outliers well
    • Modeling holidays and special events - [Link to refer](https://facebook.github.io/prophet/docs/seasonality,_holiday_effects,_and_regressors.html)
    • The model is estimated using a Bayesian approach to allow for automatic selection of the change points and other model characteristics.
    
## References of the work

- [Time Series Forecasting With Prophet In R](https://levelup.gitconnected.com/time-series-forecasting-with-prophet-in-r-a9ee81dc82e1)
- [Seasonality, Holiday Effects, And Regressors](https://facebook.github.io/prophet/docs/seasonality,_holiday_effects,_and_regressors.html)
- [Generate Quick and Accurate Time Series Forecasts using Facebook’s Prophet (with Python & R codes)](https://www.analyticsvidhya.com/blog/2018/05/generate-accurate-forecasts-facebook-prophet-python-r/)
    
## Import all the required libraries
```{r}
library(tidyverse)
library(stringr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyverse)
library(prophet)
```
```{r}
#install.packages("prophet")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Load training csv file
```{r}
train <- readr::read_csv("train.csv")
```
## Load test file 

```{r}
test <- readr::read_csv("test.csv")
```


# View few enteries of train data
```{r}
head(train)
```


## Check total Null values in the train data 
```{r}
sum(is.na(train))
```

## Check null values of each columns 
```{r}

lapply(train,function(x) { length(which(is.na(x)))})
```
> There are 2970 completely empty rows in data so we will remove those rows
> For segement 2 branch id and zone are missing, so we will first deselect those columns and then drop the null values to avoid losing 2nd segemsnt data

```{r}
raw_data = subset(train, select = -c(branch_id,zone) )
if (sum(is.na(raw_data))!= 0)
{
  data <- na.omit(raw_data)
} else
{
  data <- raw_data
}
head(data)
```

## Check unique values in segement
```{r}
unique(data$segment)
```


## Summary of the new data generated after dropping of null values 

```{r}
summary(data)
```

## Convert application date in proper date format
```{r}
data$application_date <- as.Date(data$application_date, "%d-%m-%Y") 
```
```{r}
head(data$application_date)
```


## Convert application date in proper date format
```{r}
train$application_date <- as.Date(train$application_date, "%d-%m-%Y") 
test$application_date <- as.Date(test$application_date, "%d-%m-%Y") 
```

```{r}
head(data$application_date)
```

## Split the dataset into two seperate dataframe based on 2 segement 

## Splitting test data 
```{r}
test_s1 <- test[which(test$segment == 1),]

test_s2 <- test[which(test$segment == 2),]
```

```{r}
class(test$application_date) 
```
## Splitting train data based on 2 different segemnt 
```{r}
# splitting dataframe based on segments 1 and 2
train_s1 <- train[which(train$segment == 1),]%>%
  group_by(application_date) 

train_s2 <- train[which(train$segment == 2),]%>%
  group_by(application_date) 
```
```{r}
train_s1
```

```{r}
unique(train_s1$segment)
```

## Subsetting the train and test data for model building and prediction

```{r}
df_1 = subset(train_s1, select = c(application_date,no_of_applicants) )
df_2 = subset(train_s2, select = c(application_date,no_of_applicants) )
```
```{r}
test_s1 = subset(test_s1, select = c(application_date) )
test_s2 = subset(test_s2, select = c(application_date) )
```
```{r}
head(test_s1)
```

## Building 4 test and train dataset with variable names “ds” and “y” representing the application_date and no of application counts to accommodate prophet

```{r}
test_s1 <- mutate (
test_s1,
ds = application_date,  # Create new ds column from date using mutate

)

test_s2 <- mutate (
test_s2,
ds = application_date,  # Create new ds column from date using mutate

)
test_s1 = subset(test_s1, select = c(ds) )
test_s2 = subset(test_s2, select = c(ds) )
```
```{r}
df_1 <- mutate (
df_1,
ds = application_date,  # Create new ds column from date using mutate
y = no_of_applicants   # Create new y column from value using mutate
)

df_2 <- mutate (
df_2,
ds = application_date,  # Create new ds column from date using mutate
y = no_of_applicants   # Create new y column from value using mutate
)
```


# Prophet Model building
```{r}
# Model building
m = prophet(df_1)
future = make_future_dataframe(m, periods = 7)
forecast = predict(m, future)
```

```{r}
tail(forecast)

```

```{r}
plot(m, forecast)

```

```{r}
prophet_plot_components(m, forecast)
```

## Prophet Model Building for segement 1 data with weekly and daily sesonality set as true for better forecast plot
```{r}
# Model building
m = prophet(df_1,weekly.seasonality=TRUE, daily.seasonality=TRUE)
future = make_future_dataframe(m, periods = 30)
forecast = predict(m, future)
```

```{r}
prophet_plot_components(m, forecast)
```

## predicting the values for test dataset for segment 1 and storing in y_s1
```{r}
y_s1 <- predict(m, test_s1)
```
```{r}
nrow(test_s1)
```


## Adding predicted values for in test dataset where segement is 1 for no. of applicants
```{r}
test[test$segment==1, 'no_of_applicants'] <- y_s1$yhat
```


## Prophet Model Building for segement 2 data with weekly and daily sesonality set as true for better forecast plot

```{r}
# Model building
m_2 = prophet(df_2,weekly.seasonality=TRUE, daily.seasonality=TRUE)
future_2 = make_future_dataframe(m_2, periods = 30, include_history = TRUE)
forecast_2 = predict(m_2, future_2)
```


## Plotting the forcast values 
```{r}
prophet_plot_components(m_2, forecast_2)
```


## Summary of the model
```{r}
summary(m_2)
```


## Predicting values for in test dataset where segement is 2 for no. of applicants
```{r}
y_s2 <- predict(m_2, test_s2)
```

## length of test data 
```{r}
nrow(test)
```

## Adding predicted values for in test dataset where segement is 2 for no. of applicants
```{r}
test[test$segment==2, 'no_of_applicants'] <- y_s2$yhat
```

## Saving the test data with predicted values as output_files.csv
```{r}
write.csv(test, "output_files.csv")
```
















